---
title: "Building Product-Level Distribution Dataset"
subtitle: "Mapping 120 GLORIA Products to Sectoral Distributions via Eora26 Crosswalk"
author: "Felipe Esteves"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  fig.width = 10,
  fig.height = 6
)
```

# Overview

This script transforms **sectoral distribution data** (26 Eora26 sectors) into
**product-level distributions** (120 GLORIA products) using a concordance matrix.

**Key Logic:**
- Each sector's value is divided equally among all products mapped to that sector
- Products can map to 1 or 2 sectors (main sector + "Other Services")
- If a product maps to 2 sectors, it receives the average of both allocations
- Final output: Value_Product sums to 1.0 per (Country, Strategy_ID)

---

# 1. Setup and Configuration

```{r libraries}
library(tidyverse)
library(readxl)
library(here)

# Suppress summarize warnings
options(dplyr.summarise.inform = FALSE)
```

```{r config}
# File paths
DATA_PATH <- "C:/Users/festeves/OneDrive - RAND Corporation/Courses/Dissertation/3rd Paper/Data"
COST_STRUCTURE_PATH <- file.path(DATA_PATH, "Cost Structure")
CROSSWALK_FILE <- file.path(DATA_PATH, "GLORIA-Eora26 - Crosswalk.xlsx")

# Countries to process
COUNTRIES <- c("BGR", "BLZ", "EGY", "LBY", "MAR", "MEX", "UGA")

# Output file
OUTPUT_FILE <- file.path(DATA_PATH, "sector_to_product_distribution_allcountries.csv")

cat("Configuration:\n")
cat("  Cost structure path:", COST_STRUCTURE_PATH, "\n")
cat("  Crosswalk file:", CROSSWALK_FILE, "\n")
cat("  Countries:", paste(COUNTRIES, collapse = ", "), "\n")
cat("  Output file:", OUTPUT_FILE, "\n\n")
```

---

# 2. Load and Validate Crosswalk Matrix

```{r load_crosswalk}
cat("="*80, "\n", sep = "")
cat("STEP 1: Loading Crosswalk Matrix\n")
cat("="*80, "\n", sep = "")

# Load crosswalk (120 products × 26 sectors in wide format)
crosswalk_wide <- read_excel(CROSSWALK_FILE, sheet = "Eora26 - GLORIA")

cat("\nCrosswalk dimensions:", nrow(crosswalk_wide), "rows ×", ncol(crosswalk_wide), "columns\n")
cat("Column names:", paste(head(names(crosswalk_wide), 10), collapse = ", "), "...\n")
cat("First few product names:\n")
print(head(crosswalk_wide$Products, 10))
```

```{r validate_crosswalk}
cat("\n--- Crosswalk Validation ---\n")

# Identify sector columns (all except 'Products' and 'Codes')
product_col <- "Products"
sector_cols <- setdiff(names(crosswalk_wide), c(product_col, "Codes", "Lfd_Nr"))

cat("Identified", length(sector_cols), "sector columns\n")
cat("Sectors:", paste(head(sector_cols, 5), collapse = ", "), "...\n")

# Check for 0s and 1s only
sector_values <- crosswalk_wide %>% select(all_of(sector_cols))
unique_vals <- unique(unlist(sector_values))
cat("Unique values in sector columns:", paste(sort(unique_vals), collapse = ", "), "\n")

if (!all(unique_vals %in% c(0, 1, NA))) {
  stop("ERROR: Crosswalk should only contain 0, 1, or NA in sector columns!")
}

# Check number of sectors per product
crosswalk_wide <- crosswalk_wide %>%
  rowwise() %>%
  mutate(n_sectors_linked = sum(c_across(all_of(sector_cols)), na.rm = TRUE)) %>%
  ungroup()

cat("\nProducts linked to sectors:\n")
print(table(crosswalk_wide$n_sectors_linked))

# Show products with multiple sector links
multi_sector <- crosswalk_wide %>%
  filter(n_sectors_linked > 1) %>%
  select(Products, n_sectors_linked)

if (nrow(multi_sector) > 0) {
  cat("\nProducts linked to multiple sectors (", nrow(multi_sector), " total):\n", sep = "")
  print(head(multi_sector, 10))
}

cat("\nValidation: OK\n")
```

---

# 3. Transform Crosswalk to Long Format

```{r crosswalk_long}
cat("\n", "="*80, "\n", sep = "")
cat("STEP 2: Transforming Crosswalk to Long Format\n")
cat("="*80, "\n", sep = "")

# Pivot longer: one row per (Product, Sector) pair where link = 1
crosswalk_long <- crosswalk_wide %>%
  select(Products, all_of(sector_cols)) %>%
  pivot_longer(
    cols = all_of(sector_cols),
    names_to = "Sector",
    values_to = "linked"
  ) %>%
  filter(linked == 1) %>%
  select(Products, Sector)

cat("Long format dimensions:", nrow(crosswalk_long), "rows (Product-Sector pairs)\n")
cat("Unique products:", n_distinct(crosswalk_long$Products), "\n")
cat("Unique sectors:", n_distinct(crosswalk_long$Sector), "\n")

# Show sample
cat("\nSample Product-Sector links:\n")
print(head(crosswalk_long, 10))
```

---

# 4. Count Products per Sector

```{r products_per_sector}
cat("\n--- Products per Sector ---\n")

products_per_sector <- crosswalk_long %>%
  group_by(Sector) %>%
  summarize(
    n_products = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(n_products))

cat("Total sectors with products:", nrow(products_per_sector), "\n\n")
print(products_per_sector)

# Add to crosswalk for easy merging
crosswalk_long <- crosswalk_long %>%
  left_join(products_per_sector, by = "Sector")
```

---

# 5. Load Sectoral Distribution Files

```{r load_distributions}
cat("\n", "="*80, "\n", sep = "")
cat("STEP 3: Loading Sectoral Distribution Files\n")
cat("="*80, "\n", sep = "")

# Function to load one country's distribution file
load_country_distribution <- function(country) {
  # File pattern: cost_str_COUNTRY.xlsx
  file_path <- file.path(COST_STRUCTURE_PATH, paste0("cost_str_", country, ".xlsx"))

  if (!file.exists(file_path)) {
    warning("File not found for country: ", country, " (", file_path, ")")
    return(NULL)
  }

  cat("  Loading:", basename(file_path), "\n")

  df <- read_excel(file_path) %>%
    mutate(Country = country) %>%
    select(Country, everything())

  return(df)
}

# Load all countries
distribution_list <- lapply(COUNTRIES, load_country_distribution)
names(distribution_list) <- COUNTRIES

# Remove NULLs (countries not found)
distribution_list <- distribution_list[!sapply(distribution_list, is.null)]

cat("\nSuccessfully loaded", length(distribution_list), "countries\n")

# Combine all distributions
distributions_all <- bind_rows(distribution_list)

cat("Combined dimensions:", nrow(distributions_all), "rows\n")
cat("Columns:", paste(names(distributions_all), collapse = ", "), "\n")
```

```{r validate_distributions}
cat("\n--- Distribution File Validation ---\n")

# Check structure
cat("Expected columns: Country, Sector, strategy_id, value\n")
cat("Actual columns:", paste(names(distributions_all), collapse = ", "), "\n")

# Column names should be: Country, Sector, strategy_id, value
# Check they exist
required_cols <- c("Sector", "strategy_id", "value")
missing_cols <- setdiff(required_cols, names(distributions_all))

if (length(missing_cols) > 0) {
  stop("ERROR: Missing required columns: ", paste(missing_cols, collapse = ", "))
}

# Convert to lowercase for consistency in code
distributions_all <- distributions_all %>%
  rename(
    country = Country,
    sector = Sector
  )

cat("\nUnique strategies:", n_distinct(distributions_all$strategy_id), "\n")
cat("Unique sectors:", n_distinct(distributions_all$sector), "\n")
cat("Unique countries:", n_distinct(distributions_all$country), "\n")

# Check that values sum to 1 per (country, strategy_id)
sum_check <- distributions_all %>%
  group_by(country, strategy_id) %>%
  summarize(total_value = sum(value), .groups = "drop")

cat("\nValue sums per (Country, Strategy_ID):\n")
print(summary(sum_check$total_value))

if (any(abs(sum_check$total_value - 1.0) > 1e-6)) {
  warning("Some (Country, Strategy_ID) combinations do not sum to 1.0!")
  bad_sums <- sum_check %>% filter(abs(total_value - 1.0) > 1e-6)
  print(head(bad_sums, 10))
}

cat("\nValidation: OK\n")
```

---

# 6. Check Sector Name Consistency

```{r check_sector_names}
cat("\n", "="*80, "\n", sep = "")
cat("STEP 4: Checking Sector Name Consistency\n")
cat("="*80, "\n", sep = "")

# Sectors in crosswalk
sectors_crosswalk <- unique(crosswalk_long$Sector)
cat("Sectors in crosswalk (", length(sectors_crosswalk), "):\n", sep = "")
print(sort(sectors_crosswalk))

# Sectors in distribution files
sectors_distribution <- unique(distributions_all$sector)
cat("\nSectors in distribution files (", length(sectors_distribution), "):\n", sep = "")
print(sort(sectors_distribution))

# Check for mismatches
in_crosswalk_not_distribution <- setdiff(sectors_crosswalk, sectors_distribution)
in_distribution_not_crosswalk <- setdiff(sectors_distribution, sectors_crosswalk)

if (length(in_crosswalk_not_distribution) > 0) {
  warning("Sectors in crosswalk but NOT in distribution files:")
  print(in_crosswalk_not_distribution)
}

if (length(in_distribution_not_crosswalk) > 0) {
  warning("Sectors in distribution files but NOT in crosswalk:")
  print(in_distribution_not_crosswalk)
}

# Standardize sector names (trim whitespace, consistent casing)
distributions_all <- distributions_all %>%
  mutate(sector = str_trim(sector))

crosswalk_long <- crosswalk_long %>%
  mutate(Sector = str_trim(Sector))

cat("\nSector name check: COMPLETE\n")
```

---

# 7. Merge and Allocate Product Values

```{r merge_allocate}
cat("\n", "="*80, "\n", sep = "")
cat("STEP 5: Merging and Allocating Product Values\n")
cat("="*80, "\n", sep = "")

# Join distributions with crosswalk
# Each sector's value will be distributed to all its products
product_distribution <- distributions_all %>%
  # Join with crosswalk on Sector
  left_join(
    crosswalk_long,
    by = c("sector" = "Sector"),
    relationship = "many-to-many"
  ) %>%
  # Keep only matched records (products that have sector links)
  filter(!is.na(Products)) %>%
  # Calculate product-level value
  # Value_Product = Value_Sector / n_products_in_sector
  mutate(
    value_sector = value,
    value_product = value_sector / n_products
  ) %>%
  # Rename for clarity
  rename(
    Country = country,
    Strategy_ID = strategy_id,
    Product = Products,
    Sector = sector,
    Value_Sector = value_sector,
    Value_Product = value_product
  ) %>%
  select(Country, Strategy_ID, Product, Sector, Value_Sector, Value_Product, n_products)

cat("Product distribution dimensions:", nrow(product_distribution), "rows\n")
cat("Unique products:", n_distinct(product_distribution$Product), "\n")
cat("Unique (Country, Strategy_ID) combinations:",
    nrow(distinct(product_distribution, Country, Strategy_ID)), "\n")

cat("\nSample output:\n")
print(head(product_distribution, 15))
```

---

# 8. Handle Multi-Sector Products

```{r handle_multi_sector}
cat("\n--- Handling Products Linked to Multiple Sectors ---\n")

# Check which products appear multiple times per (Country, Strategy_ID)
multi_sector_products <- product_distribution %>%
  group_by(Country, Strategy_ID, Product) %>%
  summarize(
    n_sectors = n(),
    sectors = paste(Sector, collapse = " + "),
    .groups = "drop"
  ) %>%
  filter(n_sectors > 1)

if (nrow(multi_sector_products) > 0) {
  cat("Found", nrow(multi_sector_products), "cases of products linked to multiple sectors\n")
  cat("Sample:\n")
  print(head(multi_sector_products, 10))

  # For products in multiple sectors, SUM their allocated values
  cat("\nSumming values for multi-sector products...\n")

  product_distribution_final <- product_distribution %>%
    group_by(Country, Strategy_ID, Product) %>%
    summarize(
      Sectors = paste(unique(Sector), collapse = " + "),
      Value_Product = sum(Value_Product),  # SUM across sectors (not average)
      .groups = "drop"
    )

} else {
  cat("No multi-sector products found - all products map to single sector\n")

  product_distribution_final <- product_distribution %>%
    select(Country, Strategy_ID, Product, Sector, Value_Product) %>%
    rename(Sectors = Sector)
}

cat("\nFinal dimensions:", nrow(product_distribution_final), "rows\n")
```

---

# 9. Validation: Check Sums

```{r validate_sums}
cat("\n", "="*80, "\n", sep = "")
cat("STEP 6: Validating Output\n")
cat("="*80, "\n", sep = "")

cat("--- Validation: Value_Product sums per (Country, Strategy_ID) ---\n")

sum_validation <- product_distribution_final %>%
  group_by(Country, Strategy_ID) %>%
  summarize(
    Total_Value_Product = sum(Value_Product),
    N_Products = n(),
    .groups = "drop"
  )

cat("\nSummary of Total_Value_Product:\n")
print(summary(sum_validation$Total_Value_Product))

# Check for deviations from 1.0
tolerance <- 1e-6
bad_sums <- sum_validation %>%
  filter(abs(Total_Value_Product - 1.0) > tolerance)

if (nrow(bad_sums) > 0) {
  warning("Found ", nrow(bad_sums), " (Country, Strategy_ID) combinations with sum != 1.0")
  cat("\nDeviant sums:\n")
  print(head(bad_sums %>% arrange(desc(abs(Total_Value_Product - 1.0))), 20))
} else {
  cat("\n✓ All (Country, Strategy_ID) combinations sum to 1.0 (within tolerance)\n")
}

# Distribution of products per strategy
cat("\nProducts per strategy:\n")
print(summary(sum_validation$N_Products))
```

---

# 10. Save Output

```{r save_output}
cat("\n", "="*80, "\n", sep = "")
cat("STEP 7: Saving Output\n")
cat("="*80, "\n", sep = "")

# Save final dataset
write_csv(product_distribution_final, OUTPUT_FILE)

cat("Saved to:", OUTPUT_FILE, "\n")
cat("Rows:", nrow(product_distribution_final), "\n")
cat("Columns:", paste(names(product_distribution_final), collapse = ", "), "\n")
```

---

# 11. Summary Statistics and Visualizations

```{r summary_stats}
cat("\n", "="*80, "\n", sep = "")
cat("SUMMARY STATISTICS\n")
cat("="*80, "\n", sep = "")

# Overall stats
cat("\nDataset overview:\n")
cat("  Total rows:", nrow(product_distribution_final), "\n")
cat("  Countries:", n_distinct(product_distribution_final$Country), "\n")
cat("  Strategies:", n_distinct(product_distribution_final$Strategy_ID), "\n")
cat("  Products:", n_distinct(product_distribution_final$Product), "\n")

# By country
cat("\nRows per country:\n")
country_counts <- product_distribution_final %>%
  count(Country, name = "n_rows")
print(country_counts)

# Value distribution
cat("\nValue_Product distribution:\n")
print(summary(product_distribution_final$Value_Product))
```

```{r visualizations, fig.width=12, fig.height=8}
# Plot 1: Distribution of Value_Product
p1 <- ggplot(product_distribution_final, aes(x = Value_Product)) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.7) +
  scale_x_log10() +
  labs(
    title = "Distribution of Value_Product (log scale)",
    x = "Value_Product",
    y = "Count"
  ) +
  theme_minimal()

print(p1)

# Plot 2: Top 20 products by average value
top_products <- product_distribution_final %>%
  group_by(Product) %>%
  summarize(Mean_Value = mean(Value_Product), .groups = "drop") %>%
  arrange(desc(Mean_Value)) %>%
  head(20)

p2 <- ggplot(top_products, aes(x = reorder(Product, Mean_Value), y = Mean_Value)) +
  geom_bar(stat = "identity", fill = "darkgreen", alpha = 0.7) +
  coord_flip() +
  labs(
    title = "Top 20 Products by Average Value_Product",
    x = "Product",
    y = "Mean Value_Product"
  ) +
  theme_minimal()

print(p2)

# Plot 3: Compare across countries for one strategy
sample_strategy <- product_distribution_final$Strategy_ID[1]

p3 <- product_distribution_final %>%
  filter(Strategy_ID == sample_strategy) %>%
  ggplot(aes(x = Country, y = Value_Product)) +
  geom_boxplot(fill = "coral", alpha = 0.7) +
  labs(
    title = paste("Value_Product Distribution Across Countries (Strategy", sample_strategy, ")"),
    x = "Country",
    y = "Value_Product"
  ) +
  theme_minimal()

print(p3)
```

---

# 12. Final Verification Report

```{r final_report}
cat("\n", "="*80, "\n", sep = "")
cat("FINAL VERIFICATION REPORT\n")
cat("="*80, "\n", sep = "")

# 1. Check product coverage
all_gloria_products <- unique(crosswalk_wide$Products)
products_in_output <- unique(product_distribution_final$Product)

missing_products <- setdiff(all_gloria_products, products_in_output)

cat("Product Coverage:\n")
cat("  Total GLORIA products:", length(all_gloria_products), "\n")
cat("  Products in output:", length(products_in_output), "\n")
cat("  Missing products:", length(missing_products), "\n")

if (length(missing_products) > 0) {
  warning("Some products are missing from output:")
  print(head(missing_products, 10))
}

# 2. Check strategy coverage
expected_strategies <- 67  # As per instructions
actual_strategies <- n_distinct(product_distribution_final$Strategy_ID)

cat("\nStrategy Coverage:\n")
cat("  Expected strategies:", expected_strategies, "\n")
cat("  Actual strategies:", actual_strategies, "\n")

# 3. Final sum check
final_sum_check <- product_distribution_final %>%
  group_by(Country, Strategy_ID) %>%
  summarize(Sum = sum(Value_Product), .groups = "drop") %>%
  mutate(Deviation = abs(Sum - 1.0))

cat("\nFinal Sum Check:\n")
cat("  Mean sum:", mean(final_sum_check$Sum), "\n")
cat("  Max deviation from 1.0:", max(final_sum_check$Deviation), "\n")
cat("  % within tolerance (1e-6):",
    round(100 * mean(final_sum_check$Deviation < 1e-6), 2), "%\n")

# 4. Example output
cat("\nExample Output (first 20 rows for one country-strategy):\n")
example <- product_distribution_final %>%
  filter(Country == COUNTRIES[1], Strategy_ID == sample_strategy) %>%
  arrange(desc(Value_Product)) %>%
  head(20)

print(example)

cat("\n", "="*80, "\n", sep = "")
cat("PROCESSING COMPLETE!\n")
cat("="*80, "\n", sep = "")
cat("\nOutput file:", OUTPUT_FILE, "\n")
```

---

# Session Info

```{r session_info}
sessionInfo()
```
