---
title: "GLORIA Products to Sectoral Distribution Allocation"
author: "Dataset Construction"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readxl)
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(knitr)
```

# Overview

This document creates a dataset connecting **120 GLORIA products** to sectoral distribution values across **7 countries** (BGR, BLZ, EGY, LBY, MAR, MEX, UGA) using a crosswalk matrix.

**Key transformation**: Allocate each sector's value evenly among all products linked to that sector.

---

# Step A: Load Data

## A.1 Load Country Distribution Files

```{r load-country-files}
# Define data paths
data_path <- "C:/Users/festeves/OneDrive - RAND Corporation/Courses/Dissertation/3rd Paper/Data/Cost Structure"
crosswalk_path <- "C:/Users/festeves/OneDrive - RAND Corporation/Courses/Dissertation/3rd Paper/Data"

# Country codes
countries <- c("BGR", "BLZ", "EGY", "LBY", "MAR", "MEX", "UGA")

# Load all country files
country_data_list <- list()
for (country in countries) {
  file_path <- file.path(data_path, paste0("cost_str_", country, ".xlsx"))
  cat("Loading:", country, "\n")

  df <- read_excel(file_path)
  df$Country <- country

  # Validate structure
  expected_cols <- c("Sector", "strategy_id", "value")
  if (!all(expected_cols %in% names(df))) {
    stop(paste("Missing columns in", country, "file"))
  }

  country_data_list[[country]] <- df

  # Print summary
  cat("  Rows:", nrow(df),
      "| Sectors:", length(unique(df$Sector)),
      "| Strategies:", length(unique(df$strategy_id)), "\n")
}

# Combine all countries
all_countries <- bind_rows(country_data_list)
cat("\n✓ Total combined rows:", nrow(all_countries), "\n")
```

## A.2 Validate Country Data

```{r validate-country-data}
# Check that all files have same structure
validation <- all_countries %>%
  group_by(Country) %>%
  summarise(
    n_rows = n(),
    n_sectors = n_distinct(Sector),
    n_strategies = n_distinct(strategy_id),
    min_value = min(value, na.rm = TRUE),
    max_value = max(value, na.rm = TRUE)
  )

kable(validation, caption = "Country File Validation")

# Check if values sum to 1 per strategy per country
sum_check <- all_countries %>%
  group_by(Country, strategy_id) %>%
  summarise(total_value = sum(value), .groups = "drop") %>%
  mutate(valid = abs(total_value - 1.0) < 1e-10)

if (any(!sum_check$valid)) {
  warning("Some strategy_id sums do not equal 1!")
  print(sum_check %>% filter(!valid))
} else {
  cat("\n✓ All strategies sum to 1.0 (within tolerance)\n")
}

# List all unique sectors
cat("\nUnique sectors (", length(unique(all_countries$Sector)), "):\n")
cat(paste(sort(unique(all_countries$Sector)), collapse = "\n"))
```

## A.3 Load Crosswalk Matrix

```{r load-crosswalk}
# Load crosswalk
crosswalk_file <- file.path(crosswalk_path, "GLORIA-Eora26 - Crosswalk.xlsx")
crosswalk_raw <- read_excel(crosswalk_file, sheet = "Eora26 - GLORIA")

cat("Crosswalk dimensions:", nrow(crosswalk_raw), "×", ncol(crosswalk_raw), "\n")
cat("Column names:", paste(names(crosswalk_raw)[1:5], collapse = ", "), "...\n")
cat("Total products:", nrow(crosswalk_raw), "\n")
```

---

# Step B: Transform Crosswalk to Long Form

```{r transform-crosswalk}
# Get product column name (first column) and sector columns (remaining)
product_col <- names(crosswalk_raw)[1]
sector_cols <- names(crosswalk_raw)[-1]

cat("Product column:", product_col, "\n")
cat("Number of sector columns:", length(sector_cols), "\n")

# Clean product names (remove extra whitespace, standardize)
crosswalk_raw[[product_col]] <- trimws(crosswalk_raw[[product_col]])

# Check for duplicates after cleaning
n_products_before <- nrow(crosswalk_raw)
n_products_unique <- n_distinct(crosswalk_raw[[product_col]])
cat("\nProducts before cleaning:", n_products_before, "\n")
cat("Unique products after trimming:", n_products_unique, "\n")

if (n_products_before != n_products_unique) {
  warning("Found duplicate product names after trimming whitespace!")
  dupes <- crosswalk_raw[[product_col]][duplicated(crosswalk_raw[[product_col]])]
  cat("Duplicates:", paste(unique(dupes), collapse = ", "), "\n")
}

# Pivot to long format (Product-Sector pairs where value = 1)
crosswalk_long <- crosswalk_raw %>%
  pivot_longer(
    cols = all_of(sector_cols),
    names_to = "Sector",
    values_to = "linked"
  ) %>%
  filter(linked == 1) %>%
  select(-linked) %>%
  rename(Product = all_of(product_col)) %>%
  mutate(Product = trimws(Product))  # Clean product names again

cat("\n✓ Long format crosswalk:", nrow(crosswalk_long), "Product-Sector pairs\n")
head(crosswalk_long, 10) %>% kable(caption = "First 10 Product-Sector Links")
```

## Step B.5: Verify Multi-Sector Products

```{r verify-multi-sector}
# Count how many sectors each product is linked to
multi_sector_check <- crosswalk_long %>%
  group_by(Product) %>%
  summarise(n_sectors = n(), sectors = paste(Sector, collapse = " + ")) %>%
  arrange(desc(n_sectors))

cat("\nDistribution of products by number of linked sectors:\n")
print(table(multi_sector_check$n_sectors))

cat("\nProducts linked to multiple sectors:\n")
multi_sector_products <- multi_sector_check %>% filter(n_sectors > 1)
print(multi_sector_products)

if (nrow(multi_sector_products) == 0) {
  cat("\n✓ No products linked to multiple sectors\n")
} else {
  cat("\n⚠ Found", nrow(multi_sector_products), "product(s) linked to multiple sectors\n")
}
```

---

# Step C: Compute Products-Per-Sector

```{r products-per-sector}
# Count number of products per sector
products_per_sector <- crosswalk_long %>%
  group_by(Sector) %>%
  summarise(n_products_in_sector = n()) %>%
  arrange(desc(n_products_in_sector))

cat("Products per sector:\n")
print(products_per_sector)

# Visualization
ggplot(products_per_sector, aes(x = reorder(Sector, n_products_in_sector),
                                 y = n_products_in_sector)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Number of GLORIA Products per Eora26 Sector",
       x = "Sector", y = "Number of Products") +
  theme_minimal()
```

## Step C.5: Check for Sectors with Zero Products

```{r check-zero-products}
# Get all sectors from country files
all_sectors_in_data <- unique(all_countries$Sector)

# Check if any sector has zero products
sectors_with_products <- unique(products_per_sector$Sector)
sectors_without_products <- setdiff(all_sectors_in_data, sectors_with_products)

if (length(sectors_without_products) > 0) {
  cat("⚠ WARNING: Sectors with NO products assigned:\n")
  print(sectors_without_products)
} else {
  cat("✓ All sectors in the data have at least one product\n")
}

# Add zero check
if (any(products_per_sector$n_products_in_sector == 0)) {
  stop("ERROR: Found sectors with zero products - would cause division by zero!")
}
```

---

# Step D: Merge and Allocate

## D.1 Join Crosswalk with Country Data

```{r merge-data}
# Merge crosswalk with country distribution data
merged_data <- all_countries %>%
  inner_join(crosswalk_long, by = "Sector") %>%
  left_join(products_per_sector, by = "Sector")

cat("Merged data rows:", nrow(merged_data), "\n")
cat("Unique combinations:\n")
cat("  Countries:", n_distinct(merged_data$Country), "\n")
cat("  Strategies:", n_distinct(merged_data$strategy_id), "\n")
cat("  Products:", n_distinct(merged_data$Product), "\n")
cat("  Sectors:", n_distinct(merged_data$Sector), "\n")
```

## D.2 Calculate Product Allocations

```{r calculate-allocations}
# Calculate Value_Product for each product
# For products linked to multiple sectors, this will create multiple rows
# which we'll sum in the next step
allocated_data <- merged_data %>%
  mutate(
    Value_Sector = value,
    Value_Product_per_sector = Value_Sector / n_products_in_sector
  ) %>%
  select(Country, strategy_id, Product, Sector,
         Value_Sector, n_products_in_sector, Value_Product_per_sector)

cat("Allocated data (before aggregation):", nrow(allocated_data), "rows\n")
head(allocated_data, 10) %>% kable(caption = "First 10 Allocated Rows")
```

## D.3 Aggregate Multi-Sector Products

```{r aggregate-multi-sector}
# For products linked to multiple sectors, sum their allocations
final_allocation <- allocated_data %>%
  group_by(Country, strategy_id, Product) %>%
  summarise(
    Sectors = paste(unique(Sector), collapse = " + "),
    n_sectors_linked = n_distinct(Sector),
    Value_Product = sum(Value_Product_per_sector),
    .groups = "drop"
  )

cat("\n✓ Final allocation:", nrow(final_allocation), "rows\n")
cat("  Unique products:", n_distinct(final_allocation$Product), "\n")

# DIAGNOSTIC: Check row count
expected_rows <- n_distinct(final_allocation$Country) *
                 n_distinct(final_allocation$strategy_id) *
                 n_distinct(final_allocation$Product)
actual_rows <- nrow(final_allocation)

cat("\nDIAGNOSTIC ROW COUNT CHECK:\n")
cat("  Countries:", n_distinct(final_allocation$Country), "\n")
cat("  Strategies:", n_distinct(final_allocation$strategy_id), "\n")
cat("  Products:", n_distinct(final_allocation$Product), "\n")
cat("  Expected rows:", expected_rows, "\n")
cat("  Actual rows:", actual_rows, "\n")
cat("  Difference:", actual_rows - expected_rows, "\n")

if (actual_rows != expected_rows) {
  warning("Row count mismatch! Checking for duplicates...")

  # Check for duplicate (Country, strategy_id, Product) combinations
  dup_check <- final_allocation %>%
    group_by(Country, strategy_id, Product) %>%
    summarise(n = n(), .groups = "drop") %>%
    filter(n > 1)

  if (nrow(dup_check) > 0) {
    cat("\n⚠ Found", nrow(dup_check), "duplicate (Country, strategy_id, Product) combinations:\n")
    print(head(dup_check, 20))

    # Show example
    example <- dup_check[1,]
    cat("\nExample duplicate rows:\n")
    example_rows <- final_allocation %>%
      filter(Country == example$Country,
             strategy_id == example$strategy_id,
             Product == example$Product)
    print(example_rows)
  }
}

# Show multi-sector products
multi_sector_final <- final_allocation %>%
  filter(n_sectors_linked > 1) %>%
  arrange(Country, strategy_id)

if (nrow(multi_sector_final) > 0) {
  cat("\nMulti-sector product allocations (first 10):\n")
  print(head(multi_sector_final, 10))
}
```

## D.6: Remove Any Duplicate Rows

```{r remove-duplicates}
# Just in case there are any duplicates, keep only distinct rows
rows_before <- nrow(final_allocation)

final_allocation <- final_allocation %>%
  distinct(Country, strategy_id, Product, .keep_all = TRUE)

rows_after <- nrow(final_allocation)

if (rows_before != rows_after) {
  cat("⚠ Removed", rows_before - rows_after, "duplicate rows\n")
} else {
  cat("✓ No duplicate rows found\n")
}
```

## Step D.5: Verify No Negative or NA Values

```{r verify-values}
# Check for NA values
na_check <- final_allocation %>%
  summarise(
    na_products = sum(is.na(Product)),
    na_values = sum(is.na(Value_Product))
  )

if (na_check$na_products > 0 | na_check$na_values > 0) {
  warning("Found NA values in final allocation!")
  print(na_check)
} else {
  cat("✓ No NA values found\n")
}

# Check for negative values
negative_check <- final_allocation %>%
  filter(Value_Product < 0)

if (nrow(negative_check) > 0) {
  warning("Found negative values!")
  print(negative_check)
} else {
  cat("✓ No negative values found\n")
}

# Summary statistics
cat("\nValue_Product summary:\n")
print(summary(final_allocation$Value_Product))
```

---

# Step E: Final Validation

## E.1 Verify Sums Equal 1

```{r validate-sums}
# Check that Value_Product sums to 1 per country per strategy
sum_validation <- final_allocation %>%
  group_by(Country, strategy_id) %>%
  summarise(
    total_value = sum(Value_Product),
    difference_from_1 = abs(total_value - 1.0),
    valid = difference_from_1 < 1e-10,
    .groups = "drop"
  )

# Overall validation
valid_count <- sum(sum_validation$valid)
total_count <- nrow(sum_validation)

cat("\nValidation results:\n")
cat("  Valid sums:", valid_count, "/", total_count,
    "(", round(100 * valid_count / total_count, 2), "%)\n")

if (valid_count < total_count) {
  cat("\n⚠ WARNING: Some sums do not equal 1.0\n")
  invalid_sums <- sum_validation %>%
    filter(!valid) %>%
    arrange(desc(difference_from_1))
  print(head(invalid_sums, 20))
} else {
  cat("\n✓ All strategy sums equal 1.0 (within tolerance 1e-10)\n")
}

# Histogram of differences
ggplot(sum_validation, aes(x = difference_from_1)) +
  geom_histogram(bins = 50, fill = "steelblue") +
  scale_x_log10() +
  labs(title = "Distribution of Sum Deviations from 1.0",
       x = "Absolute Difference from 1.0 (log scale)",
       y = "Count") +
  theme_minimal()
```

## E.2 Product Coverage Check

```{r product-coverage}
# Verify all 120 products appear in output
products_in_crosswalk <- unique(crosswalk_long$Product)
products_in_output <- unique(final_allocation$Product)

cat("Products in crosswalk:", length(products_in_crosswalk), "\n")
cat("Products in output:", length(products_in_output), "\n")

missing_products <- setdiff(products_in_crosswalk, products_in_output)
if (length(missing_products) > 0) {
  cat("\n⚠ Missing products:\n")
  print(missing_products)
} else {
  cat("\n✓ All products from crosswalk appear in output\n")
}
```

## E.3 Strategy Coverage Check

```{r strategy-coverage}
# Check strategies per country
strategy_summary <- final_allocation %>%
  group_by(Country) %>%
  summarise(
    n_strategies = n_distinct(strategy_id),
    n_products = n_distinct(Product),
    n_rows = n()
  )

kable(strategy_summary, caption = "Strategy and Product Coverage by Country")

# Expected: 7 countries × 67 strategies × 120 products = ~56,280 rows
expected_rows <- 7 * 67 * 120
cat("\nExpected rows:", expected_rows, "\n")
cat("Actual rows:", nrow(final_allocation), "\n")
```

---

# Step F: Save Output

```{r save-output}
# Prepare final output with clean column names
output_data <- final_allocation %>%
  select(Country, strategy_id, Product, Sectors, Value_Product) %>%
  arrange(Country, strategy_id, Product)

# Save to CSV
output_file <- "sector_to_product_distribution_allcountries.csv"
write.csv(output_data, output_file, row.names = FALSE)

cat("\n✓ Saved output to:", output_file, "\n")
cat("  Total rows:", nrow(output_data), "\n")

# Show sample of output
cat("\nFirst 20 rows of output:\n")
head(output_data, 20) %>% kable()

cat("\nLast 20 rows of output:\n")
tail(output_data, 20) %>% kable()
```

---

# Summary Statistics

```{r summary-stats}
# Overall statistics
cat("=== FINAL DATASET SUMMARY ===\n\n")
cat("Dimensions:", nrow(output_data), "rows ×", ncol(output_data), "columns\n\n")

cat("Coverage:\n")
cat("  Countries:", n_distinct(output_data$Country), "-",
    paste(unique(output_data$Country), collapse = ", "), "\n")
cat("  Strategies:", n_distinct(output_data$strategy_id), "\n")
cat("  Products:", n_distinct(output_data$Product), "\n\n")

cat("Value_Product statistics:\n")
print(summary(output_data$Value_Product))

# Distribution by country
country_stats <- output_data %>%
  group_by(Country) %>%
  summarise(
    n_rows = n(),
    mean_value = mean(Value_Product),
    median_value = median(Value_Product),
    min_value = min(Value_Product),
    max_value = max(Value_Product)
  )

cat("\n")
kable(country_stats, caption = "Value Distribution by Country", digits = 6)
```

---

# Diagnostic Plots

```{r diagnostic-plots, fig.width=10, fig.height=6}
# Value distribution
ggplot(output_data, aes(x = Value_Product)) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.7) +
  facet_wrap(~ Country, ncol = 2) +
  labs(title = "Distribution of Product Values by Country",
       x = "Value_Product", y = "Count") +
  theme_minimal()

# Box plot by country
ggplot(output_data, aes(x = Country, y = Value_Product, fill = Country)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(title = "Product Value Distribution by Country (log scale)",
       y = "Value_Product (log scale)") +
  theme_minimal() +
  theme(legend.position = "none")
```

---

# Session Info

```{r session-info}
sessionInfo()
```
