---
title: "Crosswalk Eora26 - GLORIA (SSP)"
author: "Fernando Esteves"
date: "2026-03-14"
output: html_document
---

```{r setup, include=FALSE,message=FALSE,warning=FALSE}

library(rstudioapi)
library(reshape)
library(dplyr)
library(data.table)
library(tidyverse)
library(ggplot2)
library(openxlsx)
library(readxl)
library(writexl)
library(tabr)
library(stringr)
library(summarytools)
library(gridExtra)
library(broom)
library(modelsummary)
library(rlang)
library(tidyxl)
library(unpivotr)
library(purrr)
library(knitr)

# Working directory: go up 3 levels from temp folder
wd <- dirname(dirname(rstudioapi::getActiveDocumentContext()$path))
knitr::opts_knit$set(root.dir= wd, include=FALSE, echo = TRUE)

st_options(descr.silent = TRUE)
```

## Crosswalk: Eora26 to GLORIA MRIO dataset

This document develops a new crosswalk linking the SSP cost structure to the GLORIA MRIO dataset.

A previous version established correspondence between SSP transformation costs and the Eora26‑sector classification.

Building on that work, the present script remaps those same SSP cost items to the GLORIA framework of 120 products, producing an updated structure for use in GLORIA‑based analyses and model integration.

In this script, we focus on **7 countries** (BGR, BLZ, EGY, LBY, MAR, MEX, UGA).


# 1. Generate the Cost Structure of Transformations for selected countries

```{r transformations,include=TRUE,message=FALSE}

# Cost Structure - Transformations ----

countries <- c("BGR","BLZ","EGY","LBY","MAR","MEX","UGA")

transf_sheets <- c("Synthetic_ind_energy","AGRC","CCSQ","EN","ENTC","FGTV","FRST","INEN","LNDU","LSMM","LVST","IPPU","SCOE","TRNS","TRWW","WALI","WASO")

for(i in 1:7){

for(k in 1:17){
  
  eval(parse(text=paste0("cost_str_",countries[i],"_",transf_sheets[k]," <- read_excel('Data/ATTRIBUTE_STRATEGY_transformations_vectors.xlsx',sheet ='",transf_sheets[k],"',n_max=26)")))
  
}

# All transformations in a single data frame

eval(parse(text=paste0("cost_str_",countries[i],"<- merge(cost_str_",countries[i],"_AGRC,cost_str_",countries[i],"_CCSQ,all.x = T,by='Sector')")))
}


for(i in 1:7){
for(m in 4:17){

eval(parse(text = paste0("cost_str_",countries[i]," <- merge(cost_str_",countries[i],",cost_str_",countries[i],"_",transf_sheets[m],",all.x = T,by='Sector')")))

}}


for(i in 1:7){

eval(parse(text = paste0("rm('costr_str_",countries[i],"_AGRC','costr_str_",countries[i],"_CCSQ','costr_str_",countries[i],"_EN','costr_str_",countries[i],"_ENTC','costr_str_",countries[i],"_FGTV','costr_str_",countries[i],"_FRST','costr_str_",countries[i],"_INEN','costr_str_",countries[i],"_IPPU','costr_str_",countries[i],"_LNDU','costr_str_",countries[i],"_LSMM','costr_str_",countries[i],"_LVST','costr_str_",countries[i],"_PFLO','costr_str_",countries[i],"_SCOE','costr_str_",countries[i],"_SOIL','costr_str_",countries[i],"_TRDE','costr_str_",countries[i],"_TRNS','costr_str_",countries[i],"_TRWW','costr_str_",countries[i],"_WALI','costr_str_",countries[i],"_WASO')")))

eval(parse(text = paste0("cost_str_",countries[i]," <- cost_str_",countries[i]," %>%  mutate(across(2:ncol(cost_str_",countries[i],"), function(x) ifelse(is.na(x),0,x)))")))

eval(parse(text = paste0("synthetic_ind_",countries[i]," <- cost_str_",countries[i],"_Synthetic_ind_energy"))) # Consolidating synthetic sectors

}

# BAU Electricity Generation (IEA data) 

bau_electricity <- read.csv("Data/electricity-prod-source-stacked.csv")
bau_electricity <- bau_electricity %>% filter(Code %in% c("BGR","BLZ","EGY","LBY","MAR","MEX","UGA"), Year %in% 2019)

bau_electricity <- setorder(bau_electricity,Code,-Year)

bau_electricity <- bau_electricity %>% mutate(order= sequence(rle(Code)$length)) %>% filter(order==1) %>% select(-order) %>%
                  mutate(across(4:ncol(bau_electricity), function(x) ifelse(is.na(x),0,x))) %>%
                  mutate(total_prod = rowSums(.[4:12]),
                  Bioenergy= Electricity.from.bioenergy..TWh./total_prod,Solar=Electricity.from.solar..TWh./total_prod,
                  Wind=Electricity.from.wind..TWh./total_prod,Hydro=Electricity.from.hydro..TWh./total_prod,
                  Nuclear=Electricity.from.nuclear..TWh./total_prod,
                  "Oil and Gas"=(Electricity.from.oil..TWh.+Electricity.from.gas..TWh.)/total_prod,
                  Coal=Electricity.from.coal..TWh./total_prod,Geothermal=1-Bioenergy-Solar-Wind-Hydro-Nuclear-`Oil and Gas`-Coal) %>%
                mutate(across(where(is.numeric), function(x) round(x,digits = 2))) %>%
                  select(Code,Bioenergy,Solar,Wind,Hydro,Nuclear,`Oil and Gas`,Coal,Geothermal)

# Calculating each country BAU Electricity SS ---- 

# "Renewable Energy" shares are taken from IEA's report "Energy Technology Perspectives 2023" (Global net-zero scenario)

for(i in 1:7){

eval(parse(text = paste0("bau_electricity_",countries[i]," <- bau_electricity %>% filter(Code=='",countries[i],"')")))

eval(parse(text = paste0("synthetic_ind_",countries[i]," <- synthetic_ind_",countries[i]," %>%    
  
  mutate(BAU_Electricity_",countries[i],"=Bioenergy*abs(bau_electricity_",countries[i],"$Bioenergy)+Solar*abs(bau_electricity_",countries[i],"$Solar)+
  
  Wind*abs(bau_electricity_",countries[i],"$Wind)+Hydro*abs(bau_electricity_",countries[i],"$Hydro)+Nuclear*abs(bau_electricity_",countries[i],"$Nuclear)+`Oil and Gas`*abs(bau_electricity_",countries[i],"$`Oil and Gas`)+
  
  Coal*abs(bau_electricity_",countries[i],"$Coal)+
  
  Geothermal*abs(bau_electricity_",countries[i],"$Geothermal),
  
  Renewable_Energy=0.2*Wind+Solar*0.2+Bioenergy*0.15+Hydro*0.1+`Green Hydrogen`*0.08+`Oil and Gas`*0.13+Coal*0.02+Nuclear*0.12,
  
  'Energy Efficiency'=0.5*Weatherization+0.25*`Industrial EE`+0.25*`Grid Upgrades`,
  
  'Fossil Fuels'=0.5*Coal+0.5*`Oil and Gas`) %>%
  
  mutate(across(where(is.numeric), function(x) round(x,digits = 2)))")))

# Consolidating synthetic sectors ---- 

eval(parse(text = paste0("cost_str_",countries[i]," <- merge(cost_str_",countries[i],",synthetic_ind_",countries[i],",by='Sector',all.x = T)")))

eval(parse(text = paste0("cost_str_",countries[i]," <- cost_str_",countries[i]," %>%  mutate(across(where(is.numeric), function(x) round(x,digits = 2))) %>%
  mutate(`3001`=`3001`+0.12*BAU_Electricity_",countries[i],",
                                        `3002`=`3002`+0.12*Renewable_Energy,
                                        `3059`= `Energy Efficiency`,
                                        `3060`= 0.5*Renewable_Energy+0.5*`Energy Efficiency`,
                                        `3061`= 0.7*BAU_Electricity_",countries[i],"+0.3*Hydrogen,
                                        `3062`=0.7*Renewable_Energy+0.3*`Green Hydrogen`,
                                        `3003`=0.95*Renewable_Energy+0.05*BAU_Electricity_",countries[i],",
                                        `3005`=`Grid Upgrades`,
                                        `3006`=0.5*Renewable_Energy+0.5*`Grid Upgrades`,
                                        `3010`= `3010`+0.3*Renewable_Energy,
                                        `3012`= `3010`,
                                        `3015`=`3015`+0.50*BAU_Electricity_",countries[i],",
                                        `3016`= `3016`+0.50*Renewable_Energy,
                                        `3017`=`3015`,
                                        `3018`=`3016`,
                                        `3019`=`Industrial EE`,
                                        `3020`=0.51*Renewable_Energy+0.51*`Industrial EE`,
                                        `3021`=`Industrial EE`,
                                        `3022`=0.51*Renewable_Energy+0.51*`Industrial EE`,
                                        `4002`=`4002`+0.21*BAU_Electricity_",countries[i],"+0.08*Coal,
                                        `3025`=`3025`+0.1*BAU_Electricity_",countries[i],",
                                        `3026`=`3026`+0.1*Renewable_Energy,
                                        `3028`=`3027`*0.8+0.2*Renewable_Energy,
                                        `3029`=BAU_Electricity_",countries[i],",
                                        `3030`=Renewable_Energy,
                                        `3035`=`3035`+0.1*BAU_Electricity_",countries[i],",
                                        `3036`=`3036`+0.1*Renewable_Energy,
                                        `3039`=`3039`+0.16*BAU_Electricity_",countries[i],",
                                        `3040`=`3040`+0.16*Renewable_Energy,
                                        `3041`=`3041`+0.44*Hydrogen,
                                        `3042`=`3042`+0.44*`Green Hydrogen`,
                                        `3043`=`3039`,
                                        `3044`=`3040`,
                                        `2014`= `2014`+0.103*BAU_Electricity_",countries[i],") %>% 
        
        select(-Bioenergy,-Solar,-Wind,-Hydro,-Nuclear,-`Oil and Gas`,-Coal,-Geothermal,-`Grid Upgrades`,-Weatherization,-Renewable_Energy,-`Energy Efficiency`,-`Fossil Fuels`,-`Industrial EE`,-Hydrogen,-`Green Hydrogen`,-`BAU_Electricity_",countries[i],"`) %>%
  
  mutate(across(where(is.numeric), function(x) round(x,digits = 2)))")))


# Scaling weights to sum 1

eval(parse(text = paste0("cost_str_",countries[i]," <- data.table(cost_str_",countries[i],")")))
eval(parse(text = paste0("cost_str_",countries[i]," <- melt.data.table(cost_str_",countries[i],",id.vars = 'Sector')")))
eval(parse(text = paste0("cost_str_",countries[i]," <- cost_str_",countries[i]," %>% rename(strategy_id=variable)")))

eval(parse(text = paste0("sum_check <- cost_str_",countries[i]," %>% group_by(strategy_id) %>% mutate(weight_sum=sum(value)) %>% slice_max(value,with_ties = F) %>% mutate(value = ifelse(value != 1,value+(1-weight_sum),value)) %>% select(-weight_sum)"))) 

eval(parse(text = paste0("cost_str_",countries[i]," <- merge(cost_str_",countries[i],",sum_check,by=c('Sector','strategy_id'),all.x = T)")))
eval(parse(text = paste0("cost_str_",countries[i]," <- cost_str_",countries[i]," %>% mutate(value.x = ifelse(!is.na(value.y),value.y,value.x)) %>% rename(value=value.x) %>% select(-value.y)")))
eval(parse(text = paste0("cost_str_",countries[i]," <- setorder(cost_str_",countries[i],",strategy_id,Sector)")))

}


rm(list = setdiff(ls(),c("wd","cost_str_LBY","cost_str_BGR","cost_str_BLZ","cost_str_EGY","cost_str_MAR","cost_str_MEX","cost_str_UGA")))

cost <- c("cost_str_LBY", "cost_str_BGR", "cost_str_BLZ","cost_str_EGY", 
          "cost_str_MAR", "cost_str_MEX","cost_str_UGA")

export_dir <- "Data/Cost Structure"

for(i in seq_along(cost)){

  obj_name <- cost[i]
  obj <- get(obj_name)
  
  file_name <- paste0(obj_name,".xlsx")
  file_path <- file.path(export_dir,file_name)
  
openxlsx::write.xlsx(obj, file = file_path)
}

```

# 2. Building dataset crosswalking cost structure files to 120 GLORIA products

## A.1 Load Country Distribution Files

```{r load-country-files}
# Define data paths
data_path <- "Data/Cost Structure"
crosswalk_path <- "Data"

# Country codes
countries <- c("BGR", "BLZ", "EGY", "LBY", "MAR", "MEX", "UGA")

# Load all country files
country_data_list <- list()
for (country in countries) {
  file_path <- file.path(data_path, paste0("cost_str_", country, ".xlsx"))
  cat("Loading:", country, "\n")

  df <- read_excel(file_path)
  df$Country <- country

  # Validate structure
  expected_cols <- c("Sector", "strategy_id", "value")
  if (!all(expected_cols %in% names(df))) {
    stop(paste("Missing columns in", country, "file"))
  }

  country_data_list[[country]] <- df

  # Print summary
  cat("  Rows:", nrow(df),
      "| Sectors:", length(unique(df$Sector)),
      "| Strategies:", length(unique(df$strategy_id)), "\n")
}

# Combine all countries
all_countries <- bind_rows(country_data_list)
cat("\n✓ Total combined rows:", nrow(all_countries), "\n")
```

## A.2 Validate Country Data

```{r validate-country-data}
# Check that all files have same structure
validation <- all_countries %>%
  group_by(Country) %>%
  summarise(
    n_rows = n(),
    n_sectors = n_distinct(Sector),
    n_strategies = n_distinct(strategy_id),
    min_value = min(value, na.rm = TRUE),
    max_value = max(value, na.rm = TRUE)
  )

kable(validation, caption = "Country File Validation")

# Check if values sum to 1 per strategy per country
sum_check <- all_countries %>%
  group_by(Country, strategy_id) %>%
  summarise(total_value = sum(value), .groups = "drop") %>%
  mutate(valid = abs(total_value - 1.0) < 1e-10)

if (any(!sum_check$valid)) {
  warning("Some strategy_id sums do not equal 1!")
  print(sum_check %>% filter(!valid))
} else {
  cat("\n✓ All strategies sum to 1.0 (within tolerance)\n")
}

# List all unique sectors
cat("\nUnique sectors (", length(unique(all_countries$Sector)), "):\n")
cat(paste(sort(unique(all_countries$Sector)), collapse = "\n"))
```

## A.3 Load Crosswalk Matrix

```{r load-crosswalk}
# Load crosswalk
crosswalk_file <- file.path(crosswalk_path, "GLORIA-Eora26 - Crosswalk.xlsx")
crosswalk_raw <- read_excel(crosswalk_file, sheet = "Eora26 - GLORIA")

cat("Crosswalk dimensions:", nrow(crosswalk_raw), "×", ncol(crosswalk_raw), "\n")
cat("Column names:", paste(names(crosswalk_raw)[1:5], collapse = ", "), "...\n")
cat("Total products:", nrow(crosswalk_raw), "\n")
```


### Step B: Transform Crosswalk to Long Form

```{r transform-crosswalk}

# Get product column name (first column) and sector columns (remaining)

product_col <- names(crosswalk_raw)[1]
sector_cols <- names(crosswalk_raw)[-1]

cat("Product column:", product_col, "\n")
cat("Number of sector columns:", length(sector_cols), "\n")

# Pivot to long format (Product-Sector pairs where value = 1)
crosswalk_long <- crosswalk_raw %>%
  pivot_longer(
    cols = all_of(sector_cols),
    names_to = "Sector",
    values_to = "linked"
  ) %>%
  filter(linked == 1) %>%
  select(-linked) %>%
  rename(Product = all_of(product_col))

cat("\n✓ Long format crosswalk:", nrow(crosswalk_long), "Product-Sector pairs\n")
head(crosswalk_long, 10) %>% kable(caption = "First 10 Product-Sector Links")
```

### Step B.5: Verify Multi-Sector Products

```{r verify-multi-sector}
# Count how many sectors each product is linked to
multi_sector_check <- crosswalk_long %>%
  group_by(Product) %>%
  summarise(n_sectors = n(), sectors = paste(Sector, collapse = " + ")) %>%
  arrange(desc(n_sectors))

cat("\nDistribution of products by number of linked sectors:\n")
print(table(multi_sector_check$n_sectors))

cat("\nProducts linked to multiple sectors:\n")
multi_sector_products <- multi_sector_check %>% filter(n_sectors > 1)
print(multi_sector_products)

if (nrow(multi_sector_products) == 0) {
  cat("\n✓ No products linked to multiple sectors\n")
} else {
  cat("\n⚠ Found", nrow(multi_sector_products), "product(s) linked to multiple sectors\n")
}
```

### Step C: Compute Products-Per-Sector

```{r products-per-sector}
# Count number of products per sector
products_per_sector <- crosswalk_long %>%
  group_by(Sector) %>%
  summarise(n_products_in_sector = n()) %>%
  arrange(desc(n_products_in_sector))

cat("Products per sector:\n")
print(products_per_sector)

# Visualization
ggplot(products_per_sector, aes(x = reorder(Sector, n_products_in_sector),
                                 y = n_products_in_sector)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Number of GLORIA Products per Eora26 Sector",
       x = "Sector", y = "Number of Products") +
  theme_minimal()
```

### Step C.5: Check for Sectors with Zero Products

```{r check-zero-products}
# Get all sectors from country files
all_sectors_in_data <- unique(all_countries$Sector)

# Check if any sector has zero products
sectors_with_products <- unique(products_per_sector$Sector)
sectors_without_products <- setdiff(all_sectors_in_data, sectors_with_products)

if (length(sectors_without_products) > 0) {
  cat("⚠ WARNING: Sectors with NO products assigned:\n")
  print(sectors_without_products)
} else {
  cat("✓ All sectors in the data have at least one product\n")
}

# Add zero check
if (any(products_per_sector$n_products_in_sector == 0)) {
  stop("ERROR: Found sectors with zero products - would cause division by zero!")
}
```

#### Step D: Merge and Allocate

##### D.1 Join Crosswalk with Country Data

```{r merge-data}
# Merge crosswalk with country distribution data
merged_data <- all_countries %>%
  left_join(crosswalk_long, by = "Sector") %>%
  left_join(products_per_sector, by = "Sector")

cat("Merged data rows:", nrow(merged_data), "\n")
cat("Unique combinations:\n")
cat("  Countries:", n_distinct(merged_data$Country), "\n")
cat("  Strategies:", n_distinct(merged_data$strategy_id), "\n")
cat("  Products:", n_distinct(merged_data$Product), "\n")
cat("  Sectors:", n_distinct(merged_data$Sector), "\n")
```

##### D.2 Calculate Product Allocations

```{r calculate-allocations}
# Calculate Value_Product for each product
# For products linked to multiple sectors, this will create multiple rows
# which we'll sum in the next step
allocated_data <- merged_data %>%
  mutate(
    Value_Sector = value,
    Value_Product_per_sector = Value_Sector / n_products_in_sector
  ) %>%
  select(Country, strategy_id, Product, Sector,
         Value_Sector, n_products_in_sector, Value_Product_per_sector)

cat("Allocated data (before aggregation):", nrow(allocated_data), "rows\n")
head(allocated_data, 10) %>% kable(caption = "First 10 Allocated Rows")
```

#### D.3 Aggregate Multi-Sector Products

```{r aggregate-multi-sector}
# For products linked to multiple sectors, sum their allocations
final_allocation <- allocated_data %>%
  group_by(Country, strategy_id, Product) %>%
  summarise(
    Sectors = paste(unique(Sector), collapse = " + "),
    n_sectors_linked = n_distinct(Sector),
    Value_Product = sum(Value_Product_per_sector),
    .groups = "drop"
  ) %>%
  
  mutate(Value_Product=ifelse(is.na(Value_Product),0,Value_Product))

cat("\n✓ Final allocation:", nrow(final_allocation), "rows\n")
cat("  Unique products:", n_distinct(final_allocation$Product), "\n")

# Show multi-sector products
multi_sector_final <- final_allocation %>%
  filter(n_sectors_linked > 1) %>%
  arrange(Country, strategy_id)

if (nrow(multi_sector_final) > 0) {
  cat("\nMulti-sector product allocations (first 10):\n")
  print(head(multi_sector_final, 10))
}

# Prepare final output as demand vectors to run GLORIA ($1M allocated to each strategy in each country) ----

product_number <- read.xlsx("Data/Gloria - Product list number.xlsx")

output_data <- final_allocation %>%
  select(Country, strategy_id, Product, Value_Product) %>%
  filter(!is.na(Product)) %>%
  left_join(product_number,by=c("Product"="Sector_names")) %>%
  arrange(Country, strategy_id, Product) %>%
  rename("Country ISO*"=Country,"Sector investing"=Product,"Sector investing code*"=Lfd_Nr,
         "Value*"=Value_Product) %>%
  mutate(`Value*`=`Value*`*1000000,`Type*`="abs-b",
         Country=ifelse(`Country ISO*`=="BGR","Bulgaria",
                        ifelse(`Country ISO*`=="BLZ","Belize",
                               ifelse(`Country ISO*`=="EGY","Egypt",
                                      ifelse(`Country ISO*`=="UGA","Uganda",
                                             ifelse(`Country ISO*`=="LBY","Libya",
                                                    ifelse(`Country ISO*`=="MAR","Morocco",
                                                           ifelse(`Country ISO*`=="MEX","Mexico",NA)))))))) %>%
  relocate(Country,.before = `Country ISO*`) %>% relocate(`Sector investing code*`,.after=`Sector investing`)
  
# Exporting files as to be used by MINDSET ----

export_dir <- "Mindset - WB/MINDSET_module-main/MINDSET_module-main/GLORIA_template/Scenarios"
if (!dir.exists(export_dir)) dir.create(export_dir, recursive = TRUE)

# Get unique combinations of Country, Country ISO*, and strategy_id
unique_pairs <- output_data %>%
  distinct(`Country ISO*`, Country, strategy_id)

for (i in seq_len(nrow(unique_pairs))) {

  iso    <- unique_pairs$`Country ISO*`[i]
  ctry   <- unique_pairs$Country[i]
  strat  <- unique_pairs$strategy_id[i]

  # Filter the dataset and drop the strategy_id column
  subset_df <- output_data %>%
    filter(`Country ISO*` == iso, strategy_id == strat) %>%
    select(-strategy_id)

  # Build file name, cleaning invalid characters if needed
  file_name <- paste0("Strategy_", strat, "_", iso, ".xlsx")
  file_path <- file.path(export_dir, file_name)

  # Export file
  openxlsx::write.xlsx(subset_df, file = file_path,sheetName="Investment by")

  cat("✓ Exported strategy", strat, "for", ctry, "(", iso, ") →", file_name, "\n")
}

rm(list = setdiff(ls(),c("wd","output_data")))


```

#### Step D.5: Verify No Negative or NA Values

```{r verify-values}
# Check for NA values
na_check <- final_allocation %>%
  summarise(
    na_products = sum(is.na(Product)),
    na_values = sum(is.na(Value_Product))
  )

if (na_check$na_products > 0 | na_check$na_values > 0) {
  warning("Found NA values in final allocation!")
  print(na_check)
} else {
  cat("✓ No NA values found\n")
}

# Check for negative values
negative_check <- final_allocation %>%
  filter(Value_Product < 0)

if (nrow(negative_check) > 0) {
  warning("Found negative values!")
  print(negative_check)
} else {
  cat("✓ No negative values found\n")
}

# Summary statistics
cat("\nValue_Product summary:\n")
print(summary(final_allocation$Value_Product))
```

### Step E: Final Validation

#### E.1 Verify Sums Equal 1

```{r validate-sums}
# Check that Value_Product sums to 1 per country per strategy
sum_validation <- final_allocation %>%
  group_by(Country, strategy_id) %>%
  summarise(
    total_value = sum(Value_Product),
    difference_from_1 = abs(total_value - 1.0),
    valid = difference_from_1 < 1e-10,
    .groups = "drop"
  )

# Overall validation
valid_count <- sum(sum_validation$valid)
total_count <- nrow(sum_validation)

cat("\nValidation results:\n")
cat("  Valid sums:", valid_count, "/", total_count,
    "(", round(100 * valid_count / total_count, 2), "%)\n")

if (valid_count < total_count) {
  cat("\n⚠ WARNING: Some sums do not equal 1.0\n")
  invalid_sums <- sum_validation %>%
    filter(!valid) %>%
    arrange(desc(difference_from_1))
  print(head(invalid_sums, 20))
} else {
  cat("\n✓ All strategy sums equal 1.0 (within tolerance 1e-10)\n")
}

# Histogram of differences
ggplot(sum_validation, aes(x = difference_from_1)) +
  geom_histogram(bins = 50, fill = "steelblue") +
  scale_x_log10() +
  labs(title = "Distribution of Sum Deviations from 1.0",
       x = "Absolute Difference from 1.0 (log scale)",
       y = "Count") +
  theme_minimal()
```

#### E.2 Product Coverage Check

```{r product-coverage}
# Verify all 120 products appear in output
products_in_crosswalk <- unique(crosswalk_long$Product)
products_in_output <- unique(final_allocation$Product)

cat("Products in crosswalk:", length(products_in_crosswalk), "\n")
cat("Products in output:", length(products_in_output), "\n")

missing_products <- setdiff(products_in_crosswalk, products_in_output)
if (length(missing_products) > 0) {
  cat("\n⚠ Missing products:\n")
  print(missing_products)
} else {
  cat("\n✓ All products from crosswalk appear in output\n")
}
```

#### E.3 Strategy Coverage Check

```{r strategy-coverage}
# Check strategies per country
strategy_summary <- final_allocation %>%
  group_by(Country) %>%
  summarise(
    n_strategies = n_distinct(strategy_id),
    n_products = n_distinct(Product),
    n_rows = n()
  )

kable(strategy_summary, caption = "Strategy and Product Coverage by Country")

# Expected: 7 countries × 67 strategies × 120 products = ~56,280 rows
expected_rows <- 7 * 67 * 120
cat("\nExpected rows:", expected_rows, "\n")
cat("Actual rows:", nrow(final_allocation), "\n")
```

# Summary Statistics

```{r summary-stats}
# Overall statistics
cat("=== FINAL DATASET SUMMARY ===\n\n")
cat("Dimensions:", nrow(output_data), "rows ×", ncol(output_data), "columns\n\n")

cat("Coverage:\n")
cat("  Countries:", n_distinct(output_data$Country), "-",
    paste(unique(output_data$Country), collapse = ", "), "\n")
cat("  Strategies:", n_distinct(output_data$strategy_id), "\n")
cat("  Products:", n_distinct(output_data$Product), "\n\n")

cat("Value_Product statistics:\n")
print(summary(output_data$Value_Product))

# Distribution by country
country_stats <- output_data %>%
  group_by(Country) %>%
  summarise(
    n_rows = n(),
    mean_value = mean(Value_Product),
    median_value = median(Value_Product),
    min_value = min(Value_Product),
    max_value = max(Value_Product)
  )

cat("\n")
kable(country_stats, caption = "Value Distribution by Country", digits = 6)
```

---

# Diagnostic Plots

```{r diagnostic-plots, fig.width=10, fig.height=6}
# Value distribution
ggplot(output_data, aes(x = Value_Product)) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.7) +
  facet_wrap(~ Country, ncol = 2) +
  labs(title = "Distribution of Product Values by Country",
       x = "Value_Product", y = "Count") +
  theme_minimal()

# Box plot by country
ggplot(output_data, aes(x = Country, y = Value_Product, fill = Country)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(title = "Product Value Distribution by Country (log scale)",
       y = "Value_Product (log scale)") +
  theme_minimal() +
  theme(legend.position = "none")
```
